---
title: "US Border Patrol Apprehensions"
author: "Yin-Chia Huang, Francesco Re, Emma Yue Wu (Group 6)"
date: "2/11/2018"
subtitle: "MA415 Spring 2018, Assignment 3"
runtime: shiny
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)

#Set up for shiny
# import 2017 data without sector names
library(readr)
bp17 <- read_csv("PB Apprehensions 2017.csv")[1:10,2:13]
# transpose dataset to make sectors columns
bp17 <- t(bp17)
# import sector names and make it colnames
sectors <- read_csv("PB Apprehensions 2017.csv")[,1]
sectors <- t(sectors)
colnames(bp17) <- sectors

# import 2010 data without sector names
bp10 <- read_csv("BP Apprehensions 2010.csv")[1:10,2:13]
# transpose dataset to make sectors columns
bp10 <- t(bp10)
# import sector names and make it colnames
colnames(bp10) <- sectors
#take out any unnecessary content
as.data.frame(bp10)
bp10[,1:9]
as.data.frame(bp17)
bp17[,1:9]
app17 <- t(bp17)
app10 <- t(bp10)


#Set up for the rest of the document
#monthly apprehensions by year
msum  <- read.csv("PB monthly summaries.csv", stringsAsFactors = FALSE)
years <- msum[c(1)]
msum  <- msum[-c(1)]
rownames(msum) <- as.vector(t(years))
#msum
#monthly apprehensions by Sector
ap10 <- read.csv("BP apprehensions 2010.csv")
cities <- ap10[,1]
ap10 <- ap10[-c(1,14)]
rownames(ap10) <- cities
#ap10
SumCol10 <- colSums(Filter(is.numeric, ap10))
SC10 <- colSums(app10,na.rm=TRUE)
#monthly apprehensions by Sector
ap17 <- read.csv("edited_app17.csv")
ap17 <- ap17[-c(1,2,15)]
ap17 <- ap17[-c(10),]
rownames(ap17) <- cities
SumCol17 <- colSums(Filter(is.numeric, ap17))
SC17 <- colSums(app17,na.rm=TRUE)
#ap17


```

##Introduction

Apprehensions at the US-Mexico border have declined to near-historic lows over the last few years. The objective of this report is to give a deeper insight on this change that is been occuring. Through the analysis of the data collected by the U.S. Customs and Border Protection through the years, we intend to shed light on the general trend's behaviour of this phenomenon, focusing on how factors such as time and place have influenced the given outcome.


## Data Overview

Choose a year and sector to view the data.

```{r tabsets, echo=FALSE}

shinyApp(
# Define UI for dataset viewer app ----
ui <- fluidPage(
  
  # # App title ----
  # titlePanel("BP Apprehensions"),
  
  # Sidebar layout with input and output definitions ----
  sidebarLayout(
    
    # Sidebar panel for inputs ----
    sidebarPanel(
      
      # Input: Select a dataset ----
      selectInput("dataset", "Choose a year:",
                  choices = c("2010", "2017")),
      # Input: Select a sector ----
      selectInput("sector",
                  "Sector:",
                  choices= colnames(bp17),
                  hr()),
      
      # Include clarifying text ----
      helpText("Note: Click update button after choosing"),
      
      # Input: actionButton() to defer the rendering of output ----
      # until the user explicitly clicks the button (rather than
      # doing it immediately when inputs change). This is useful if
      # the computations required to render output are inordinately
      # time-consuming.
      actionButton("update", "Update View")
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      # Output: Header + summary of distribution ----
      h4("Summary"),
      verbatimTextOutput("summary"),
      
      # Output: Header + table of distribution ----
      h4("Bar Plot"),
      plotOutput("distPlot")
    )
    
  )
),

# Define server logic to summarize and view selected dataset ----
server <- function(input, output) {
  
  # Return the requested dataset ----
  # Note that we use eventReactive() here, which depends on
  # input$update (the action button), so that the output is only
  # updated when the user clicks the button
  datasetInput <- eventReactive(input$update, {
    switch(input$dataset,
           "2010" = bp10,
           "2017" = bp17)
  }, ignoreNULL = FALSE)
  
  # Generate a summary of the dataset ----
  output$summary <- renderPrint({
    dataset <- datasetInput()
    summary(dataset[,input$sector])
  })
  
  # Show the first "n" observations ----
  # The use of isolate() is necessary because we don't want the table
  # to update whenever input$obs changes (only when the user clicks
  # the action button)
  output$distPlot <- renderPlot({
    
    # draw the chosen year bar plot with the specified sector
    dataset <- datasetInput()
    barplot(dataset[,input$sector], 
            main=paste(input$sector, "Apprehensions in", input$dataset),
            las=2,
            ylab="Number of People")
   })
  
  },
  options = list(height = 600)
)
```


## Timeline and Projections

```{r, echo=FALSE}
#Comparison total apprehensions by month 
CmbColS <- rbind((SC10),(SC17))
colours <- c("orange","red")
barplot(CmbColS,
        beside=TRUE, 
        las=2, 
        main = "Apprehensions by month",
        cex.lab =1.5,
        col = colours )
legend("topright",
       c("Apprehensions in 2010","Apprehensions in 2017"),
       cex=0.8,
       bty="n",
       fill=colours )

#Time-series graph of monthly sum
# read data
ts1 <- read.csv("PB monthly summaries_no title.csv")
# arrange the rows in ascending order with columns staring from Jan to Dec
ts2 <- ts1[ nrow(ts1):1, ]
ts2 <- cbind(subset.data.frame(ts2)[,4:12],subset.data.frame(ts2)[,1:3])
# change ts2 to class "ts"
ts3 <- ts(as.vector(unlist(t(ts2))),frequency=12,start=c(2000,1))
# monthly time series graph from 2000 t0 2017 
ts4 <-ts.plot(ts3, gpars=list(main="Timeline",xlab="year", ylab="Apprehensions", lty=c(1:3)))
# Box plot across months will give us a sense on seasonal effect 
boxplot(ts3~cycle(ts3), main="Distribution of Monthly Apprehensions",xlab="month", ylab="Apprehensions")
#  forecast
fit <- arima(ts3, order=c(1,0,0), list(order=c(2,1,0), period=12))
fore <- predict(fit, n.ahead=24)
#  error bounds at 95% confidence level
U <- fore$pred + 2*fore$se
L <- fore$pred - 2*fore$se
ts.plot(ts3, fore$pred, U, L, col=c(1,2,4,4), lty = c(1,1,2,2), main="Projection", xlab="month", ylab="Apprehensions")
legend("topleft", c("Actual", "Forecast", "Error Bounds (95% Confidence)"), col=c(1,2,4), lty=c(1,1,2))
# Decompostion of additive time series
timeseriescomponents <- decompose(ts3)
#timeseriescomponents
plot(timeseriescomponents)

```


## Results and Interpretations
#### 2010 max-sum sector: No.8 Tuscon
```{r, echo=FALSE}
#2010 monthly apprehensions by Sector
ap10 <- read.csv("BP apprehensions 2010.csv")
cities <- ap10[,1]
ap10 <- ap10[-c(1,14)]
rownames(ap10) <- cities
SumCol10 <- colSums(Filter(is.numeric, ap10))

#2017 monthly apprehensions by Sector
ap17 <- read.csv("edited_app17.csv")

ap17 <- ap17[-c(1,2,15)]
ap17 <- ap17[-c(10),]
rownames(ap17) <- cities
SumCol17 <- colSums(Filter(is.numeric, ap17))


#function to get t-tests for by-sector compasison
by_sector.t <- function(y){

   t.test(as.numeric(ap10[y,]), as.numeric(ap17[y,]), paired = TRUE)
}
# sector with max yearly sum in 2010 --> test on max-apprehension sector changes
by_sector.t(6)
```
t test: p= 0.072 > 0.05  
Conclusion: We cannot reject the null hypothesis that Tuscon has equal means in 2017 from 2010.  
Interpretation: Tuscon does not play a critical role in the fluctuations of apprehensions between in 2010 and in 2017.  

#### 2017 max-sum sector: No.6 Rio Grande Valley
```{r, echo=FALSE}
# sector with max yearly sum in 2017 --> test on max-apprehension sector changes
by_sector.t(8)
```
t test: p= 0.00 < 0.05  
Conclusion: We can reject the null hypothesis. RGV has significantly different means in 2010 from 2017, with 95% confidence.  
Interpretation: RGV plays a critical role in the fluctuations of apprehensions between in 2010 and in 2017.  

##### Testing the hyphotesis for which the three months with apprehension peak in one year are equal in mean of the same months in the other
#### Peak in 2010: March, April, May
```{r, echo=FALSE}
#Testing the hyphotesis for which the three most trafficated months in 2010 are equal in mean to the 
#same months in 2017
x17 <- msum[1,]
x10 <- msum[8,]
x17 <- x17[6:8]
x10 <- x10[6:8]
t.test(as.numeric(x10), as.numeric(x17), paired = TRUE)
```
t test: p= 0.014 < 0.05  
Conclusion: We can reject the null hypothesis. March - May 2010 means are significantly different from those in 2017, with 95% confidence.  
Interpretation: The peak in 2010 is significantly different in value from the same months in 2017. The test result correspond with the wide boxplot trend during those months. However, the small degrees of freedom makes the test less reliable.   

#### Peak in 2017: October, November, December
```{r, echo=FALSE}
#Testing the hyphotesis for which the three most trafficated months in 2017 are equal in mean to the 
#same months in 2010
x17 <- msum[1,]
x10 <- msum[8,]
# x17 <- x17[2:4]
# x10 <- x10[2:4]
x17 <- x17[1:3]
x10 <- x10[1:3]
t.test(as.numeric(x17), as.numeric(x10), paired = TRUE)
```
t test: p= 0.081 > 0.05  
Conclusion: We cannot reject the null hypothesis. October - December 2017 means are not significantly different from those in 2010.  
Interpretation: The peak in 2017 is not significantly different in value from the same months in 2010. The test result correspond with the narrow boxplot trend during those months. However, the small degrees of freedom makes the test less reliable. 





