---
title: "MA 415/ MA615 
        HW 3 Section 1 project"
output: ioslides_presentation
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Team members

- Yue Wu 
- Yin Chia Huang 
- Francesco Ignazio Re

## Outline

- Introduction
- Summary of the data set 
- Simple statistic tests
- Data visualization
- Conclusion

## Introduction
```{r}

```

## Summary of the data set

## Simple statistic tests
```{r echo=TRUE}

```


## Data visualization  
```{r include=FALSE}
#monthly apprehensions by year

msum  <- read.csv("PB monthly summaries.csv", stringsAsFactors = FALSE)
years <- msum[c(1)]
msum  <- msum[-c(1)]
rownames(msum) <- as.vector(t(years))
msum

#monthly apprehensions by Sector
ap10 <- read.csv("BP apprehensions 2010.csv")
cities <- ap10[,1]
ap10 <- ap10[-c(1,14)]
rownames(ap10) <- cities
ap10
SumCol10 <- colSums(Filter(is.numeric, ap10))

#monthly apprehensions by Sector
#ap17 <- read.delim("clipboard")
#write.csv(ap17, file = "edited_app17.csv")

ap17 <- read.csv("edited_app17.csv")

ap17 <- ap17[-c(1,2,15)]
ap17 <- ap17[-c(10),]
rownames(ap17) <- cities
SumCol17 <- colSums(Filter(is.numeric, ap17))
ap17


```


```{r}
# Comparison total apprehensions by month 
CmbColS <- rbind((SumCol10),(SumCol17))
colours <- c("orange","red")
barplot(CmbColS,
        beside=TRUE, 
        las=2, 
        main = "Apprehensions by month",
        cex.lab =1.5,
        col = colours )
legend("topright",
       c("Apprehensions in 2010","Apprehensions in 2017"),
       cex=0.8,
       bty="n",
       fill=colours )
```

## Data visualization 

```{r}
# Comparison total apprehensions by city

# CmbRowS <- rbind(rowSums(ap10), rowSums(ap17))
# barplot(CmbRowS,
#         beside=TRUE, 
#         las=2, main = "Apprehensions by city", 
#         cex.lab =1.5, 
#         col = colours )
# legend("topleft",
#           c("Apprehensions in 2010","Apprehensions in 2017"),
#           cex=0.8,
#           bty="n",
#           fill=colours )
```


## Time series plot grapgh - Monthly summary data
```{r include=FALSE}
# read data
ts1 <- read.csv("PB monthly summaries_no title.csv")

# arrange the rows in ascending order with columns staring from Jan to Dec
ts2 <- ts1[ nrow(ts1):1, ]
ts2 <- cbind(subset.data.frame(ts2)[,4:12],subset.data.frame(ts2)[,1:3])

# change ts2 to class "ts"
ts3 <- ts(as.vector(unlist(t(ts2))),frequency=12,start=c(2000,1))
```

```{r }

# monthly time series graph from 2000 t0 2017 
ts4 <-ts.plot(ts3, gpars=list(xlab="year", ylab="Apprehensions", lty=c(1:3)))

```


## Time series analysis

Box plot across months will give us a sense on seasonal effect 

```{r echo=FALSE}
# Box plot across months will give us a sense on seasonal effect 
boxplot(ts3~cycle(ts3))
```

## Time series analysis
- Decompostion of additive time series graph
```{r echo=FALSE}
# Decompostion of additive time series
timeseriescomponents <- decompose(ts3)
#timeseriescomponents
plot(timeseriescomponents)

```


## Time series prediction
```{r}
fit <- arima(ts3, order=c(1,0,0), list(order=c(2,1,0), period=12))
fore <- predict(fit, n.ahead=24)

#  error bounds at 95% confidence level

U <- fore$pred + 2*fore$se
L <- fore$pred - 2*fore$se
ts.plot(ts3, fore$pred, U, L, col=c(1,2,4,4), lty = c(1,1,2,2))
legend("topleft", c("Actual", "Forecast", "Error Bounds (95% Confidence)"), col=c(1,2,4), lty=c(1,1,2))


```

## Conclusion

