---
title: "MA 415/ MA615 
        HW 3 Section 1 project"
output: ioslides_presentation
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Team members

- Yue Wu 
- Yin Chia Huang 
- Francesco Ignazio Re

## Outline

- Introduction
- Summary of the data set 
- Descriptive Data Analysis
- Simple statistic tests
- Data visualization
- Conclusion

## Introduction

Apprehensions at the US-Mexico border have declined to near-historic lows over the last few years. The objective of this report is to give a deeper insight on this change that is been occuring. Through the analysis of the data collected by the U.S. Customs and Border Protection through the years, we intend to shed light on the general trend's behaviour of this phenomenon, focusing on how factors such as time and place have influenced the given outcome.


```{r include=FALSE}
# Raw Data
PB2010 <- read.csv("PB2010.csv")
PB2017 <- read.csv("PB2017.csv")
ap10 <- read.csv("BP apprehensions 2010.csv")
ap17 <- read.csv("edited_app17.csv")



```


## Over view of 2010 Apprehensions
The gray dots are Sector "Tucson", which had the most unstable number of apprehensions. And it also contributed the most apprehensions in 2010. 
```{r}
#PB2010 <- read.csv("PB2010.csv")
plot(PB2010$Month, PB2010$Apprehensions, col = PB2010$Sector)

# install.packages("tidyverse")
# library(tidyverse)
# 
# ggplot(data = PB2010) + 
#   geom_point(mapping = aes(x = Month, y = Apprehensions , color = Sector))
```

## Overview of 2017 Apprehensions
The pink dots are Sector "Rio Grande Valley" which contributed the most apprehensions in 2017. Notably,the number of apprehensions from Rio Grande Valley was unstable, especially it dramatically increased from September to October.

```{r}
#PB2017 <- read.csv("PB2017.csv")
plot(PB2017$Month, PB2017$Apprehensions, col = PB2017$Sector)

# install.packages("tidyverse")
# library(tidyverse)
# 
# ggplot(data = PB2017) + 
#   geom_point(mapping = aes(x = Month, y = Apprehensions , color = Sector))
```

## Statistic number of 2010 by month 
```{r}
#mean
mean_10_mon <- tapply(PB2010$Apprehensions, PB2010$Month, mean)
#median 
median_10_mon <- tapply(PB2010$Apprehensions, PB2010$Month, median)
# variance
var_10_mon <- tapply(PB2010$Apprehensions, PB2010$Month, var)
```

```{r}
PB10_statInfo_byMon <- data.frame(Mean = mean_10_mon, Median = median_10_mon, Variance = var_10_mon)
PB10_statInfo_byMon
```

## Statistic number of 2017 by month 
```{r}
#mean
mean_17_mon <- tapply(PB2017$Apprehensions, PB2017$Month, mean)
#median 
median_17_mon <- tapply(PB2017$Apprehensions, PB2017$Month, median)
# variance
var_17_mon <- tapply(PB2017$Apprehensions, PB2017$Month, var)
```

```{r}
PB17_statInfo_byMon <- data.frame(Mean = mean_17_mon, Median = median_17_mon, Variance = var_17_mon)
PB17_statInfo_byMon
```

## Statistic number of 2010 by sector 
 
```{r}
#mean
mean_10_sec <- tapply(PB2010$Apprehensions, PB2010$Sector, mean)
#median 
median_10_sec <- tapply(PB2010$Apprehensions, PB2010$Sector, median)
# variance
var_10_sec <- tapply(PB2010$Apprehensions, PB2010$Sector, var)
```

```{r}
PB10_statInfo_bySec <- data.frame(Mean = mean_10_sec, Median = median_10_sec, Variance = var_10_sec)
PB10_statInfo_bySec
```
## Statistic number of 2017 by sector 
```{r}
#mean
mean_17_sec <- tapply(PB2017$Apprehensions, PB2017$Sector, mean)
#median 
median_17_sec <- tapply(PB2017$Apprehensions, PB2017$Sector, median)
# variance
var_17_sec <- tapply(PB2017$Apprehensions, PB2017$Sector, var)
```

```{r}
PB17_statInfo_bySec <- data.frame(Mean = mean_17_sec, Median = median_17_sec, Variance = var_17_sec)
PB17_statInfo_bySec
```

## 2010 v.s 2017 Apprehensions by month  
The most 3 trafficated months in 2010, April, May and June, are also the ones that have witnessed the greatest decline, figuring the lowest numbers in the data collected in 2017.
While apprehensions tend to decrease during the winter months, we can see an increase from January to March from 2010 to 2017.
```{r echo=FALSE}
#ap10 <- read.csv("BP apprehensions 2010.csv")
cities <- ap10[,1]
ap10 <- ap10[-c(1,14)]
rownames(ap10) <- cities
SumCol10 <- colSums(Filter(is.numeric, ap10))
#ap17 <- read.csv("edited_app17.csv")
ap17 <- ap17[-c(1,2,15)]
rownames(ap17) <- cities
SumCol17 <- colSums(Filter(is.numeric, ap17))
CmbColS <- rbind((SumCol10),(SumCol17))
CmbRowS <- rbind(rowSums(ap10), rowSums(ap17))
colours <- c("orange","red")



#monthly apprehensions by year
msum  <- read.csv("PB monthly summaries.csv", stringsAsFactors = FALSE)
years <- msum[c(1)]
msum  <- msum[-c(1)]
rownames(msum) <- as.vector(t(years))

```

```{r echo=FALSE}
barplot(CmbColS,
        beside=TRUE, 
        las=2, 
        main = "Apprehensions by month",
        cex.lab =1.5,
        col = colours )
legend("topright",
       c("Apprehensions in 2010","Apprehensions in 2017"),
       cex=0.8,
       bty="n",
       fill=colours )
```

## 2010 v.s 2017 Apprehensions by sector
The greatest change has occured in Tuscon, the area with the highest number of apprehensions in the 2010, that observed a drop of over the 80 percent according to the data collected in 2017.
```{r echo=FALSE}
CmbRowS <- rbind(rowSums(ap10), rowSums(ap17))
barplot(CmbRowS,
        beside=TRUE, 
        las=2, main = "Apprehensions by sector", 
        cex.lab =1.5, 
        col = colours )
legend("topleft",
          c("Apprehensions in 2010","Apprehensions in 2017"),
          cex=0.8,
          bty="n",
          fill=colours )
```

## Conclusion (1)
From 2010 to 2017, the U.S. Customs and Border Pro- tection saw an overall 36 percent decrease in individuals apprehended while trying to enter the country illegally.
Taking into consideration the months and the sectors where the apprehensions respectively in 2010 and 2017 occurred, we can see that the change has been charac- terized by a major drop in certain months and certain places rather then by a constant decline everywhere.   

## Statistic Analysis(1)
### Use simple statistical tests to compare the sector with the most apprehensions for 2010 and with the sector with the most apprehensions in 2017. Has there been a change in the maximum?

---

####**2010 max-sum sector: No.8 Tuscon**
```{r echo=FALSE}
#2010 monthly apprehensions by Sector
ap10 <- read.csv("BP apprehensions 2010.csv")
cities <- ap10[,1]
ap10 <- ap10[-c(1,14)]
rownames(ap10) <- cities
SumCol10 <- colSums(Filter(is.numeric, ap10))

#2017 monthly apprehensions by Sector
ap17 <- read.csv("edited_app17.csv")

ap17 <- ap17[-c(1,2,15)]
ap17 <- ap17[-c(10),]
rownames(ap17) <- cities
SumCol17 <- colSums(Filter(is.numeric, ap17))

#function to get t-tests for by-sector compasison
by_sector.t <- function(y){

   t.test(as.numeric(ap10[y,]), as.numeric(ap17[y,]), paired = TRUE)
}
# sector with max yearly sum in 2010 --> test on max-apprehension sector changes
by_sector.t(6)
```

######**t test:** p= 0.072 > 0.05  

######**Conclusion:** We cannot reject the null hypothesis that Tuscon has equal means in 2017 from 2010.  

######**Interpretation:** Tuscon does not play a critical role in the fluctuations of apprehensions between in 2010 and in 2017.  

---

#### **2017 max-sum sector: No.6 Rio Grande Valley**
```{r echo=FALSE}
# sector with max yearly sum in 2017 --> test on max-apprehension sector changes
by_sector.t(8)
```
######**t test:** p= 0.00 < 0.05  
######**Conclusion:** We can reject the null hypothesis. RGV has significantly different means in 2010 from 2017, with 95% confidence.  
######**Interpretation:** RGV plays a critical role in the fluctuations of apprehensions between in 2010 and in 2017.  


## Statistic Analysis(2)
### Testing the hyphotesis for which the three months with apprehension peak in one year are equal in mean of the same months in the other

---

##### **Peak in 2010: March, April, May**
```{r, echo=FALSE}
#Testing the hyphotesis for which the three most trafficated months in 2010 are equal in mean to the 
#same months in 2017
x17 <- msum[1,]
x10 <- msum[8,]
x17 <- x17[6:8]
x10 <- x10[6:8]
t.test(as.numeric(x10), as.numeric(x17), paired = TRUE)
```
######**t test:** p= 0.014 < 0.05  
######**Conclusion:** We can reject the null hypothesis. March - May 2010 means are significantly different from those in 2017, with 95% confidence.  
######**Interpretation:** The peak in 2010 is significantly different in value from the same months in 2017. The test result correspond with the wide boxplot trend during those months. However, the small degrees of freedom makes the test less reliable.   

---

####**Peak in 2017: October, November, December**
```{r, echo=FALSE}
#Testing the hyphotesis for which the three most trafficated months in 2017 are equal in mean to the 
#same months in 2010
x17 <- msum[1,]
x10 <- msum[8,]
# x17 <- x17[2:4]
# x10 <- x10[2:4]
x17 <- x17[1:3]
x10 <- x10[1:3]
t.test(as.numeric(x17), as.numeric(x10), paired = TRUE)
```

**t test:** p= 0.081 > 0.05  
**Conclusion:** We cannot reject the null hypothesis. Oct - Dec 2017 means are not significantly different from those in 2010.  
**Interpretation:** The peak in 2017 is not significantly different in value from the same months in 2010. The test result correspond with the narrow boxplot trend during those months. However, the small degrees of freedom makes the test less reliable. 


## Time series grapgh 

######*The number apprehnsions dropped sharply in 2017. It was a sign of decreased traffic of illegal immigrants. This trend indicates Trump's policies and executive actions on immigration are already causing some to stop before coming.*  
```{r include=FALSE}
# read data
ts1 <- read.csv("PB monthly summaries_no title.csv")

# arrange the rows in ascending order with columns staring from Jan to Dec
ts2 <- ts1[ nrow(ts1):1, ]
ts2 <- cbind(subset.data.frame(ts2)[,4:12],subset.data.frame(ts2)[,1:3])

# change ts2 to class "ts"
ts3 <- ts(as.vector(unlist(t(ts2))),frequency=12,start=c(2000,1))
```

```{r }
# monthly time series graph from 2000 t0 2017 
ts4 <-ts.plot(ts3, gpars=list(xlab="year", ylab="Apprehensions", lty=c(1:3)))
```

## Time series analysis
Box plot across months gives us a sense on seasonal effect. We can see that the number of apprehension in March has the biggest range and highest median.

```{r echo=FALSE}
# Box plot across months will give us a sense on seasonal effect 
boxplot(ts3~cycle(ts3))
```

## Time series analysis
There is an obvious seasonal fluctuation in the time series graph. A reasonable explanation is because of the weather. The apprehensions tend to decrease during the winter months. And the decreasing trend implied that the administration's early efforts to thwart the illegal aliens problems have an effect in the long run. 
```{r echo=FALSE}
# Decompostion of additive time series
timeseriescomponents <- decompose(ts3)
#timeseriescomponents
plot(timeseriescomponents)

```


## Month plot
This graph display the time series plot for each month from 2000 to 2017. The shape of each month is very similar with different magnitudes. We need to do further research to find the cause of two peaks and troughs during this 17 years. This may explain the factor affecting apprehensions.  
```{r}
monthplot(ts3, ylab= "Apprehensions" )
```

## Time series prediction
```{r}
fit <- arima(ts3, order=c(1,0,0), list(order=c(2,1,0), period=12))
fore <- predict(fit, n.ahead=24)

#  error bounds at 95% confidence level

U <- fore$pred + 2*fore$se
L <- fore$pred - 2*fore$se
ts.plot(ts3, fore$pred, U, L, col=c(1,2,4,4), lty = c(1,1,2,2))
legend("topleft", c("Actual", "Forecast", "Error Bounds (95% Confidence)"), col=c(1,2,4), lty=c(1,1,2))

```

## Conclusion

## Reference
<br> - https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html#time-series-analysis
<br> - http://www2.hawaii.edu/~fuleky/econ427/6_Time_series_decomposition.html#1
<br> - https://rstudio-pubs-static.s3.amazonaws.com/254962_6d87741d9ad64a32b6e47d684875dde9.html
<br> - https://www.datascience.com/blog/introduction-to-forecasting-with-arima-in-r-learn-data-science-tutorials
<br> - http://www.washingtonexaminer.com/border-apprehensions-hit-17-year-low-in-march/article/2619338

