---
title: "MA 415/ MA615 
        HW 3 Section 1 project"
output: ioslides_presentation
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Team members

- Yue Wu 
- Yin Chia Huang 
- Francesco Ignazio Re

## Outline

- Introduction
- Data set Visualization & Analysis
- Statistic Analysis
- Time Series & Analysis
- Conclusion

## Introduction

Apprehensions at the US-Mexico border have declined to near-historic lows over the last few years. The objective of this report is to give a deeper insight on this change that is been occuring. Through the analysis of the data collected by the U.S. Customs and Border Protection through the years, we intend to shed light on the general trend's behaviour of this phenomenon, focusing on how factors such as time and place have influenced the given outcome.


```{r include=FALSE}
# Raw Data
PB2010 <- read.csv("PB2010.csv")
PB2017 <- read.csv("PB2017.csv")
```


## Overview of 2010 Apprehensions
The gray dots are Sector "Tucson", which had the most unstable number of apprehensions. And it also contributed the most apprehensions in 2010. 
```{r}
#PB2010 <- read.csv("PB2010.csv")

plot(PB2010$Month, PB2010$Apprehensions, type = "p",col = PB2010$Sector, main = "Apprehensions by month in 2010")
legend("topright",inset=c(0.22,0),
       c("Big Bend","Del Rio", "El Centro","El Paso","Laredo","Rio Grande Valley", "San Diego", "Tucson","Yuma"),
       cex=0.8,
       bty="n",
       fill=c("gray0","red","green","blue","cyan", "magenta","yellow","gray", "black"))


#  install.packages("tidyverse")
#  library(tidyverse)
# #
#  ggplot(data = PB2010) + 
#    geom_point(mapping = aes(x = Month, y = Apprehensions , color = Sector))
```

## Overview of 2017 Apprehensions
The pink dots are Sector "Rio Grande Valley" which contributed the most apprehensions in 2017. Notably,apprehensions dramatically increased from Sep to Oct.

```{r}
#PB2017 <- read.csv("PB2017.csv")
plot(PB2017$Month, PB2017$Apprehensions, col = PB2017$Sector, main = "Apprehensions by month in 2017")

legend("topleft",inset=c(0.05,0),
       c("Big Bend","Del Rio", "El Centro","El Paso","Laredo","Rio Grande Valley", "San Diego", "Tucson","Yuma"),
       cex=0.8,
       bty="n",
       fill=c("gray0","red","green","blue","cyan", "magenta","yellow","gray", "black"))

# install.packages("tidyverse")
# library(tidyverse)
# 
# ggplot(data = PB2017) + 
#   geom_point(mapping = aes(x = Month, y = Apprehensions , color = Sector))
```


```{r include=FALSE}
# read data for 2010 2017 comparisons
ap10 <- read.csv("BP apprehensions 2010.csv")
cities <- ap10[,1]
ap10 <- ap10[-c(1,14)]
ap10 <- cbind(ap10[4:12],ap10[1:3])
rownames(ap10) <- cities
SumCol10 <- colSums(Filter(is.numeric, ap10))

ap17 <- read.csv("edited_app17.csv")
ap17 <- ap17[-c(1,2,15)]
ap17 <- cbind(ap17[4:12],ap17[1:3])
rownames(ap17) <- cities
SumCol17 <- colSums(Filter(is.numeric, ap17))

CmbColS <- rbind((SumCol10),(SumCol17))
CmbRowS <- rbind(rowSums(ap10), rowSums(ap17))
colours <- c("orange","red")
```


## 2010 v.s 2017 Apprehensions by month  
While apprehensions tend to decrease during the winter months, we can see an obvious increase in Nov and Dec from 2010 to 2017.
```{r echo=FALSE}
barplot(CmbColS,
        beside=TRUE, 
        las=2, 
        main = "Apprehensions by month",
        cex.lab =1.5,
        col = colours )
legend("topright",
       c("Apprehensions in 2010","Apprehensions in 2017"),
       cex=0.8,
       bty="n",
       fill=colours )
```

## 2010 v.s 2017 Apprehensions by sector
The greatest change has occured in Tuscon, the area with the highest number of apprehensions in the 2010, that observed a drop of over the 80% according to the data collected in 2017.
```{r echo=FALSE}
CmbRowS <- rbind(rowSums(ap10), rowSums(ap17))
barplot(CmbRowS,
        beside=TRUE, 
        las=2, 
        main = "Apprehensions by sector", 
        cex.lab =1.5, 
        col = colours )
legend("topleft",
          c("Apprehensions in 2010","Apprehensions in 2017"),
          cex=0.8,
          bty="n",
          fill=colours )
```

## Conclusion (1)
From 2010 to 2017, the U.S. Customs and Border Pro- tection saw an overall 36 percent decrease in individuals apprehended while trying to enter the country illegally.
Taking into consideration the months and the sectors where the apprehensions respectively in 2010 and 2017 occurred, we can see that the change has been charac- terized by a major drop in certain months and certain places rather then by a constant decline everywhere.   

## Statistic Analysis (1)
###Use simple statistical tests to compare the sector with the most apprehensions for 2010 and with the sector with the most apprehensions in 2017. Has there been a change in the maximum?

--- 

#### *2010 max-sum sector: No.8 Tuscon*
```{r, echo=FALSE}
msum  <- read.csv("PB monthly summaries.csv", stringsAsFactors = FALSE)
years <- msum[c(1)]
msum  <- msum[-c(1)]

#2010 monthly apprehensions by Sector
ap10 <- read.csv("BP apprehensions 2010.csv")
cities <- ap10[,1]
ap10 <- ap10[-c(1,14)]
rownames(ap10) <- cities
SumCol10 <- colSums(Filter(is.numeric, ap10))

#2017 monthly apprehensions by Sector
ap17 <- read.csv("edited_app17.csv")

ap17 <- ap17[-c(1,2,15)]
ap17 <- ap17[-c(10),]
rownames(ap17) <- cities
SumCol17 <- colSums(Filter(is.numeric, ap17))

#function to get t-tests for by-sector compasison
by_sector.t <- function(y){

   t.test(as.numeric(ap10[y,]), as.numeric(ap17[y,]), paired = TRUE)
}
# sector with max yearly sum in 2010 --> test on max-apprehension sector changes
by_sector.t(8)
```
######**t test:** p= 0.072 > 0.05  
######**Conclusion:** We cannot reject the null hypothesis that Tuscon has equal means in 2017 from 2010.  
######**Interpretation:** Tuscon does not play a critical role in the fluctuations of apprehensions between in 2010 and in 2017.  

---

#### *2017 max-sum sector: No.6 Rio Grande Valley*
```{r, echo=FALSE}
# sector with max yearly sum in 2017 --> test on max-apprehension sector changes
by_sector.t(6)
```
######**t test:** p= 0.00 < 0.05  
######**Conclusion:** We can reject the null hypothesis. RGV has significantly different means in 2010 from 2017, with 95% confidence.  
######**Interpretation:** RGV plays a critical role in the fluctuations of apprehensions between in 2010 and in 2017.  

## Statistic Analysis (2) 
###Testing the hyphotesis for which the three months with apprehension peak in one year are equal in mean of the same months in the other. 

---

#### *Peak in 2010: March, April, May*
```{r, echo=FALSE}
x17 <- msum[1,]
x10 <- msum[8,]
x17 <- x17[6:8]
x10 <- x10[6:8]
t.test(as.numeric(x10), as.numeric(x17), paired = TRUE)
```
######**t test:** p= 0.014 < 0.05  
######**Conclusion:** We can reject the null hypothesis. March - May 2010 means are significantly different from those in 2017, with 95% confidence.  
######**Interpretation:** The peak in 2010 is significantly different in value from the same months in 2017. The test result correspond with the wide boxplot trend.    

---

#### *Peak in 2017: October, November, December*
```{r, echo=FALSE}
#Testing the hyphotesis for which the three most trafficated months in 2017 are equal in mean to the 
#same months in 2010
x17 <- msum[1,]
x10 <- msum[8,]
x17 <- x17[2:4]
x10 <- x10[2:4]
t.test(as.numeric(x17), as.numeric(x10), paired = TRUE)
```
######**t test:** p= 0.276 > 0.05  
######**Conclusion:** We cannot reject the null hypothesis. Oct - Dec 2017 means are not significantly different from those in 2010.  
######**Interpretation:** The peak in 2017 is not significantly different in value from the same months in 2010. The test result correspond with the narrow boxplot trend during those months. However, the small degrees of freedom makes the test less reliable. 

## Time series graph  
######The number apprehnsions dropped.It was a sign of decreased traffic of illegal immigrants. 
```{r include=FALSE}
# read data
ts1 <- read.csv("PB monthly summaries_no title.csv", header = FALSE)
# arrange the rows in ascending order with columns staring from Jan to Dec
ts2 <- ts1[ nrow(ts1):1, ]
ts2 <- cbind(subset.data.frame(ts2)[,4:12],subset.data.frame(ts2)[,1:3])
# change ts2 to class "ts"
ts3 <- ts(as.vector(unlist(t(ts2))),frequency=12,start=c(2000,1))
```

```{r include=FALSE}
# Create monthly mean data for second time series line

## calculate the annaul means 
ts3.1 <- as.vector(unlist(t(ts2)))

for (i in 1:18){
  if (i == 1){
    from <- i
    to <- 12
    means <- c(mean(ts3.1[from :to]))
  } 
  else{
    from <- 1 + to
    to <- i*12
  }
  means <- append(means, (mean(ts3.1[from :to])))
}

## delete the duplicated means in 2010
means <- means[-1]
## repete the annual means 12 times for one year
means1 <- rep(means,each=12)
## make the data into matrix to add headers and row names
means2 <- matrix(means1, nrow = 18, ncol = 12, byrow=TRUE) 
dimnames(means2) <- list(
  2000: 2017,
  c("Jan", "Feb", "March", "April", "May", "June", "July", "Aug", "Sep", "Oct", "Nov", "Dec"))
## change data type ti class ts
means.ts <- ts(as.vector(unlist(t(means2))),frequency=12,start=c(2000,1))

```

```{r echo=FALSE}
## Plot the time Series Graph (including two lines)
ts4 <- ts.plot(ts3,means.ts, gpars=list(xlab="year", ylab="Apprehensions", lty=c(1:3)), main = "Time Series from 2010 to 2017")
## change the color
lines(means.ts, lwd=2, lty=1, col = rgb( 0, .7, .9, .5) )
## add text labels
text(0.5+seq(from = 2000, to = 2017, by = 1),means[1:18]+7000,labels=2000:2017, cex=0.6, col = rgb(.9,  0, .7, .5) , font=2)
## add legend
legend('topright', col=c(rgb(0,0,0),  rgb( 0, .7, .9, .5)), lty=1, lwd=2, 
       legend=c("Monthly apprehensions", "Annual averages"), bg='white')  

```

## Time series analysis (1)
Box plot across months will give us a sense on seasonal effect. We can see that the number of apprehension in March has the biggest range and highest median.

```{r echo=FALSE}
# Box plot across months will give us a sense on seasonal effect 
boxplot(ts3~cycle(ts3),ylab= "Apprehensions", xlab= "Month", main="Monthly Box plot " )
```


##Time series analysis (2)
######There is an obvious seasonal fluctuation in the time series graph. A reasonable explanation is weather. The apprehensions tend to decrease during the winter months.But the decreasing trend implied that the administration's early efforts to thwart the illegal aliens problems have an effect in the long run. 
```{r echo=FALSE}
# Decompostion of multiplicative time series
timeseriescomponents <- decompose(ts3, type="multiplicative")

#timeseriescomponents
plot(timeseriescomponents)

```


## Time series analysis (3)
Monthly plot graph displays the time series plot for each month from 2000 to 2017. The horizontal lines indicate the means for each month. The shape of each month is very similar with different magnitudes.  
```{r}
monthplot(ts3, ylab= "Apprehensions", xlab= "Month", main="Monthly plot with monthly mean")
```

## Time series prediction
Using ARIMA Model to predict future apprehensions
```{r}
fit <- arima(ts3, order=c(1,0,0), list(order=c(2,1,0), period=12))
fore <- predict(fit, n.ahead=24)

#  error bounds at 95% confidence level
U <- fore$pred + 2*fore$se
L <- fore$pred - 2*fore$se
ts.plot(ts3, fore$pred, U, L, col=c(1,2,4,4), lty = c(1,1,2,2), main="Future Apprenhensions Prediction")
legend("topleft", c("Actual", "Forecast", "Error Bounds (95% Confidence)"), col=c(1,2,4), lty=c(1,1,2))

```

## Conclusion

## Reference
<br> - https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html#time-series-analysis
<br> - http://www2.hawaii.edu/~fuleky/econ427/6_Time_series_decomposition.html#1
<br> - https://rstudio-pubs-static.s3.amazonaws.com/254962_6d87741d9ad64a32b6e47d684875dde9.html
<br> - https://www.datascience.com/blog/introduction-to-forecasting-with-arima-in-r-learn-data-science-tutorials
<br> - http://www.washingtonexaminer.com/border-apprehensions-hit-17-year-low-in-march/article/2619338
<br> - http://ucanalytics.com/blogs/arima-models-manufacturing-case-study-example-part-3/

