---
title: "MA 415/ MA615 
        HW 3 Section 1 project"
output: ioslides_presentation
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Team members

- Yue Wu 
- Yin Chia Huang 
- Francesco Ignazio Re

## Outline

- Introduction
- Summary of the data set 
- Simple statistic tests
- Data visualization
- Conclusion

## Introduction

Apprehensions at the US-Mexico border have declined to near-historic lows over the last few years. The objective of this report is to give a deeper insight on this change that is been occuring. Through the analysis of the data collected by the U.S. Customs and Border Protection through the years, we intend to shed light on the general trend's behaviour of this phenomenon, focusing on how factors such as time and place have influenced the given outcome.

## Summary of 2010 Apprehensions
```{r}
PB2010 <- read.csv("PB2010.csv")
plot(PB2010$Month, PB2010$Apprehensions, col = PB2010$Sector)

# install.packages("tidyverse")
# library(tidyverse)
# 
# ggplot(data = PB2010) + 
#   geom_point(mapping = aes(x = Month, y = Apprehensions , color = Sector))
```

## Summary of 2017 Apprehensions
```{r}
PB2017 <- read.csv("PB2017.csv")
plot(PB2017$Month, PB2017$Apprehensions, col = PB2017$Sector)

# install.packages("tidyverse")
# library(tidyverse)
# 
# ggplot(data = PB2017) + 
#   geom_point(mapping = aes(x = Month, y = Apprehensions , color = Sector))
```

## Statistic number of 2010 by month 
mean 
```{r}
tapply(PB2010$Apprehensions, PB2010$Month, mean)
```
median  
```{r}
tapply(PB2010$Apprehensions, PB2010$Month, median)
```
variance 
```{r}
tapply(PB2010$Apprehensions, PB2010$Month, var)
```


## Statistic number of 2017 by month 
mean 
```{r}
tapply(PB2017$Apprehensions, PB2017$Month, mean)
```
median  
```{r}
tapply(PB2017$Apprehensions, PB2017$Month, median)
```
variance 
```{r}
tapply(PB2017$Apprehensions, PB2017$Month, var)
```

## Statistic number of 2010 by sector 
mean 
```{r}
tapply(PB2010$Apprehensions, PB2010$Sector, mean)
```
median 
```{r}
tapply(PB2010$Apprehensions, PB2010$Sector, median)
```
variance
```{r}
tapply(PB2010$Apprehensions, PB2010$Sector, var)
```

## Statistic number of 2017 by sector 
mean 
```{r}
tapply(PB2017$Apprehensions, PB2017$Sector, mean)
```
median 
```{r}
tapply(PB2017$Apprehensions, PB2017$Sector, median)
```
variance
```{r}
tapply(PB2017$Apprehensions, PB2017$Sector, var)
```

## Data visualization  
```{r include=FALSE}
#monthly apprehensions by year

msum  <- read.csv("PB monthly summaries.csv", stringsAsFactors = FALSE)
years <- msum[c(1)]
msum  <- msum[-c(1)]
rownames(msum) <- as.vector(t(years))
msum

#monthly apprehensions by Sector
ap10 <- read.csv("BP apprehensions 2010.csv")
cities <- ap10[,1]
ap10 <- ap10[-c(1,14)]
rownames(ap10) <- cities
ap10
SumCol10 <- colSums(Filter(is.numeric, ap10))

#monthly apprehensions by Sector
#ap17 <- read.delim("clipboard")
#write.csv(ap17, file = "edited_app17.csv")

ap17 <- read.csv("edited_app17.csv")

ap17 <- ap17[-c(1,2,15)]
ap17 <- ap17[-c(10),]
rownames(ap17) <- cities
SumCol17 <- colSums(Filter(is.numeric, ap17))
ap17


```


```{r}
# Comparison total apprehensions by month 
CmbColS <- rbind((SumCol10),(SumCol17))
colours <- c("orange","red")
barplot(CmbColS,
        beside=TRUE, 
        las=2, 
        main = "Apprehensions by month",
        cex.lab =1.5,
        col = colours )
legend("topright",
       c("Apprehensions in 2010","Apprehensions in 2017"),
       cex=0.8,
       bty="n",
       fill=colours )
```

## Data visualization 

```{r}
# Comparison total apprehensions by city

# CmbRowS <- rbind(rowSums(ap10), rowSums(ap17))
# barplot(CmbRowS,
#         beside=TRUE, 
#         las=2, main = "Apprehensions by city", 
#         cex.lab =1.5, 
#         col = colours )
# legend("topleft",
#           c("Apprehensions in 2010","Apprehensions in 2017"),
#           cex=0.8,
#           bty="n",
#           fill=colours )
```


## Time series plot grapgh - Monthly summary data
```{r include=FALSE}
# read data
ts1 <- read.csv("PB monthly summaries_no title.csv")

# arrange the rows in ascending order with columns staring from Jan to Dec
ts2 <- ts1[ nrow(ts1):1, ]
ts2 <- cbind(subset.data.frame(ts2)[,4:12],subset.data.frame(ts2)[,1:3])

# change ts2 to class "ts"
ts3 <- ts(as.vector(unlist(t(ts2))),frequency=12,start=c(2000,1))
```

```{r }

# monthly time series graph from 2000 t0 2017 
ts4 <-ts.plot(ts3, gpars=list(xlab="year", ylab="Apprehensions", lty=c(1:3)))

```


## Time series analysis

Box plot across months will give us a sense on seasonal effect 

```{r echo=FALSE}
# Box plot across months will give us a sense on seasonal effect 
boxplot(ts3~cycle(ts3))
```

## Time series analysis
- Decompostion of additive time series graph
```{r echo=FALSE}
# Decompostion of additive time series
timeseriescomponents <- decompose(ts3)
#timeseriescomponents
plot(timeseriescomponents)

```


## Time series prediction
```{r}
fit <- arima(ts3, order=c(1,0,0), list(order=c(2,1,0), period=12))
fore <- predict(fit, n.ahead=24)

#  error bounds at 95% confidence level

U <- fore$pred + 2*fore$se
L <- fore$pred - 2*fore$se
ts.plot(ts3, fore$pred, U, L, col=c(1,2,4,4), lty = c(1,1,2,2))
legend("topleft", c("Actual", "Forecast", "Error Bounds (95% Confidence)"), col=c(1,2,4), lty=c(1,1,2))

```

## Conclusion


